An Embedding Framework for Consistent Polyhedral Surrogates

### 2021-08-05 poly-embed-journal/poly-embed-journal.tex
meeting notes

% Convex surrogate losses are a central building block in machine learning for finite prediction problems such as classification and structured prediction tasks.
A growing body of work seeks to design and analyze convex surrogates for particular target loss functions, and more broadly, understand the best empirical risk minimization bounds that can found for a surrogate, for which consistency is a necessary condition.
For example, recent work has developed tools to bound the prediction dimension of the surrogate, meaning the dimension of the range of the surrogate hypothesis~\cite{frongillo2015elicitation,  ramaswamy2016convex}.
Yet in some cases these bounds are far from tight, such as for \emph{abstain loss} (classification with an abstain option)~\citep{bartlett2008classification,yuan2010classification,ramaswamy2016convex,ramaswamy2018consistent,zhang2018reject}.
Furthermore, the kinds of strategies available for constructing surrogates, and their relative power, are not well understood.

@@

We assume that a given discrete prediction problem, such as classification, is given in the form of a discrete \emph{target loss} $\ell:\R\to\reals^\Y_+$, which maps a report (prediction) $r$ from a finite set $\R$ to the vector of loss values $\ell(r) = (\ell(r)_y)_{y\in\Y}$ for each possible outcome $y\in\Y$.
%We will assume throughout that the given discrete loss is \emph{non-redundant}, meaning every report is uniquely optimal (minimizes expected loss) for some distribution $p\in\simplex$.
Similarly, surrogate losses will be written $L:\reals^d\to\reals^\Y_+$, typically with reports written $u\in\reals^d$.
%\btw{Cut generic loss -- from neurips 2019 lol}%(A generic loss will be written $L:\R\to\reals^\Y$.)
We write the corresponding expected loss when $Y \sim p$ as $\inprod{p}{\ell(r)}$ and $\inprod{p}{L(u)}$.
The \emph{Bayes risk} of a loss $L:\reals^d\to\reals^\Y_+$ is the function $\risk{L}:\simplex\to\reals_+$ given by $\risk{L}(p) := \inf_{u\in\reals^d} \inprod{p}{L(u)}$; naturally for discrete losses we write $\risk{\ell}$ with the infimum over $\R$.
When restricting the domain of a loss $L$ from $\R$ to $\R'$, we write $L|_{\R'}$.

@@

For finite properties (those with $|\R|<\infty$) and discrete losses, we will use lowercase notation $\gamma$ and $\ell$, respectively, with reports $r\in\R$; for surrogate properties and losses we use $\Gamma$ and $L$, with reports $u\in\reals^d$.
% For finite properties (those with $|\R|<\infty$) and discrete losses, we will use lowercase notation $\gamma:\simplex\toto\R$ and $\ell:\R\toto\reals^\Y_+$, respectively, with reports $r\in\R$; for surrogate properties and losses we use $\Gamma:\simplex\toto\reals^d$ and $L:\reals^d\toto\reals^\Y_+$, with reports $u\in\reals^d$.
For general properties and losses, we will also use $\Gamma$ and $L$, as above.
\bo{This slightly clashes with statement above in 2.1 that $L$ means a surrogate loss.}

@@

In fact, it is not always necessary to map \emph{all} reports to such points.
It turns out to often be convenient to allow $\ell$ to have reports that are ``redundant'' in some sense (explored in \S~\ref{sec:min-rep-sets}), initially explored by~\citet{wang2020weston} under the name \emph{embedding cardinality}.
%\bo{perhaps mention and cite \citet{wang2020weston}}
Because of this, a key notion to define an embedding is that of a \emph{representative set}: a set of reports $\Sc$ such that, for all label distributions, at least one report $r\in\Sc$ minimizes expected loss.
\begin{definition}[Representative set]
  Let $\Gamma:\simplex\toto\R$.
  We say $\Sc \subseteq \R$ is \emph{representative for $\Gamma$} if we have $\Gamma(p) \cap \Sc \neq \emptyset$ for all $p\in \simplex$.
  We further say $\Sc$ is a \emph{minimum representative set} if it has the smallest cardinality among all representative sets.
  Given a loss $L:\R\to\reals^\Y_+$, we say $\Sc$ is a (minimum) representative set for $L$ if it is a (minimum) representative set for $\prop L$.
  % We say $\Sc \subseteq \R$ is \emph{representative for $L$} if we have $\prop{L}(p) \cap \Sc \neq \emptyset$ for all $p\in \simplex$.
  % We further say $\Sc$ is a \emph{minimum representative set} if it has the smallest cardinality among all representative sets.
\end{definition}
\bo{Removed a paragraph here and folded it into the above. It's good to explain why we need representative sets -- i.e. the first sentence was great. But I don't think we should go into any details of various ways in which reports can be redundant. It doesn't matter, we can forward ref. Instead though, I think we should give some intuition for why it will be convenient.}

@@

To illustrate the idea of embedding, let us examine hinge loss in detail as a surrogate for 0-1 loss for binary classification.
Recall that we have $\R = \Y = \{-1, +1\}$, with $\hinge(u)_y = (1 - uy)_+$ and $\ellzo(r)_y := \Ind{r\neq y}$, typically with link function $\psi(u) = \sgn(u)$.
We will see that hinge loss embeds (2 times) 0-1 loss, via the embedding $\varphi(r) = r$.
For condition (i), it is straightforward to check that $\hinge(r)_y = 2\ellzo(r)_y$ for all $r,y\in\{-1,1\}$.
% First, consider that
% \begin{align*}
% 	\ellzo(-1) = (0,1) && \ellzo(1) = (1,0)\\
% 	L_{hinge}(-1) = (0,2) && L_{hinge}(1) = (2,0)\\
% \end{align*}
% Therefore, $L_{hinge}$ is twice $\ellzo$ for each $r \in \R$.
For condition (ii), let us compute the property each loss elicits, i.e., the set of optimal reports for each $p$:
\[
\prop{\ellzo}(p) = \begin{cases}
1 & p_1 > 1/2 \\
\{-1,1\} & p_1 = 1/2\\
-1 & p_1 < 1/2
\end{cases}
\qquad
\prop{L_{hinge}}(p) = \begin{cases}
[1,\infty) & p_1 = 1\\
1 & p_1 \in (1/2,1) \\
[-1,1] & p_1 = 1/2\\
-1& p_1 \in (0, 1/2)\\
(-\infty, -1]& p_1 = 0
\end{cases}~.
\]
% Now, take the embedding $\varphi$ to be the identity.  % := r$ if $r \in [-1,1]$, and $\sgn(r)$ otherwise.
In particular, we see that $-1 \in \prop{\ellzo}(p) \iff p_1 \in [0, 1/2] \iff -1 \in \prop{\hinge}(p)$, and $1 \in \prop{\ellzo}(p) \iff p_1 \in [1/2,1] \iff 1 \in \prop{\hinge}(p)$.
With both conditions of Definition~\ref{def:loss-embed} satisfied, we conclude that $\hinge$ embeds $2\ellzo$.

\begin{figure}
	\begin{minipage}{0.3\linewidth}
	\centering
	\includegraphics[width=0.95\linewidth]{figs/0-1-br.pdf}
%		\caption{Bayes risk of the 0-1 loss.}
%		\label{fig:0-1-br}
	\end{minipage}
	\hfill
	\begin{minipage}{0.3\linewidth}
	\centering		\includegraphics[width=0.95\linewidth]{figs/hinge-br.pdf}
%		\caption{Bayes risk of hinge loss.}
%		\label{fig:hinge-br}
	\end{minipage}
	\hfill
	\begin{minipage}{0.3\linewidth}
	\centering
	\includegraphics[width=0.95\linewidth]{figs/logistic-br.pdf}
%	\caption{Bayes risk of the logistic loss.}
%	\label{fig:logistic-br}
\end{minipage}
\caption{Bayes risks of 0-1, hinge, and logistic losses, respectively, in binary classification setting. \jessie{Move this down to when we actually discuss risks matching?}}
\label{fig:bayes-risks-01}
\end{figure}

@@

In this section, we establish a tight relationship between the technique of embedding and the use of polyhedral (piecewise-linear convex) surrogate losses.
We defer the question of when such surrogates are consistent to \S~\ref{sec:calibration}.

@@

A first observation is that a loss restricted to some representative set $\Sc$ elicits the same property restricted to $\Sc$.
As a consequence, restricting to representative sets preserves the Bayes risk.
We will use these observations throughout.
\begin{lemma}\label{lem:loss-restrict}
  Let $L:\R\to\reals^\Y_+$ elicit $\Gamma$, and let $\Sc\subseteq\R$ be representative for $L$.
  % such that $\Gamma(p) \cap \Sc \neq \emptyset$ for all $p\in\simplex$.
  Then $L|_\Sc$ elicits $\gamma:\simplex\toto\Sc$ defined by $\gamma(p) = \Gamma(p)\cap \Sc$.
  Moreover, $\risk{L}=\risk{L|_\Sc}$.
\end{lemma}
\botodo{I didn't see us define $L|_\Sc$ anywhere yet! Although it's briefly mentioned in the intro.}\raft{Yeah, maybe recall the def here?}
\botodo{Notation in the lemma is confusing because $L$ is supposed to be a surrogate with report space $\reals^d$ and $\R$ is supposed to be the report space for $\ell$.}\raft{We should discuss this point}
\begin{proof}
  Let $p\in\simplex$ be fixed throughout.
  First let $r \in \gamma(p) = \Gamma(p) \cap \Sc$.
  Then $r \in \Gamma(p) = \argmin_{u\in\R} \inprod{p}{L(u)}$, so as $r\in\Sc$ we have in particular $r \in \argmin_{u\in\Sc} \inprod{p}{L(u)}$.
  For the other direction, suppose $r \in \argmin_{u\in\Sc} \inprod{p}{L(u)}$.
  As $\Sc$ is representative for $L$, we must have some $s \in \Gamma(p) \cap \Sc$.
  On the one hand, $s\in\Gamma(p) = \argmin_{u\in\R} \inprod{p}{L(u)}$.
  On the other, as $s \in \Sc$, we certainly have $s \in \argmin_{u\in\Sc} \inprod{p}{L(u)}$.
  But now we must have $\inprod{p}{L(r)} = \inprod{p}{L(s)}$, and thus $r \in \argmin_{u\in\R} \inprod{p}{L(u)} = \Gamma(p)$ as well.
  We now see $r \in \Gamma(p) \cap \Sc$.
  Finally, the equality of the Bayes risks $\min_{u\in\R} \inprod{p}{L(u)} = \min_{u\in\Sc} \inprod{p}{L(u)}$ follows immediately by the above, as $\emptyset \neq \Gamma(p)\cap\Sc \subseteq \Gamma(p)$ for all $p\in\simplex$.
\end{proof}

@@

@@

@@

In light of this negative result, together with the fact that every polyhedral surrogate embeds some target loss (Theorem~\ref{thm:poly-embeds-discrete}), we now turn to the question: what target problem, then, \emph{is} the Lov\'asz hinge consistent for?
We show in \S~\ref{app:lovasz} that the answer is a variant of abstain loss where one can abstain on a set of labels $B\subseteq[k]$.
Recall that the symmetric difference is defined by $S\triangle T = (S\setminus T) \cup (T\setminus S)$.
\begin{theorem}\label{thm:lovasz-embed-main}
	Let $\hat\ell^f:\hat\R\times\Y\to\reals_+$, where $\hat\R = \{(A,B) \in 2^{[k]} \times 2^{[k]}: A\cap B=\emptyset\}$, be given by
	\begin{equation}
	\hat\ell^f((A,B),S) = f(A\triangle S\setminus B) + f(A\triangle S\cup B)~.
	\end{equation}
	Then the Lov\'asz hinge $L^f$ embeds $\hat\ell^f$.
  Moreover, the embedding is tight whenever $f$ is strictly submodular.
\end{theorem}
\raf{Something about how this could actually be a useful target and surrogate, provided you actually link from the surrogate to the target and not ignore $B$.}

>>>>

% Convex surrogate losses are a central building block in machine learning for finite prediction problems such as classification and structured prediction tasks.
A growing body of work seeks to design and analyze convex surrogates for particular target loss functions, and more broadly, understand the best empirical risk minimization bounds that can be found for a surrogate, for which consistency is a necessary condition.
For example, recent work has developed tools to bound the prediction dimension of the surrogate, meaning the dimension of the range of the surrogate hypothesis~\cite{frongillo2015elicitation,  ramaswamy2016convex}.
Yet in some cases these bounds are far from tight, such as for \emph{abstain loss} (classification with an abstain option)~\citep{bartlett2008classification,yuan2010classification,ramaswamy2016convex,ramaswamy2018consistent,zhang2018reject}.
Furthermore, the kinds of strategies available for constructing surrogates, and their relative power, are not well understood.

@@

\bo{Suggest: ``...loss function $L:\R\to\reals^\Y_+$...'' then transition to target and surrogate}
\raf{Moved: `` which maps a report (prediction) $r$ from a finite set $\R$ to the vector of loss values $L(r) = (L(r)_y)_{y\in\Y}$ for each possible outcome $y\in\Y$''}
We write the corresponding expected loss when $Y \sim p$ as $\inprod{p}{\ell(r)}$ and $\inprod{p}{L(u)}$.
The \emph{Bayes risk} of a loss $L:\reals^d\to\reals^\Y_+$ is the function $\risk{L}:\simplex\to\reals_+$ given by $\risk{L}(p) := \inf_{u\in\reals^d} \inprod{p}{L(u)}$; naturally for discrete losses we write $\risk{\ell}$ with the infimum over $\R$.
When restricting the domain of a loss $L$ from $\R$ to $\R'$, we write $L|_{\R'}$.

We assume that a given discrete prediction problem, such as classification, is given in the form of a discrete \emph{target loss}, which we denote $\ell:\R\to\reals^\Y_+$, where we assume $\R$ is a finite set.
%We will assume throughout that the given discrete loss is \emph{non-redundant}, meaning every report is uniquely optimal (minimizes expected loss) for some distribution $p\in\simplex$.
Similarly, surrogate losses will be written $L:\reals^d\to\reals^\Y_+$, typically with reports written $u\in\reals^d$.
%\btw{Cut generic loss -- from neurips 2019 lol}%(A generic loss will be written $L:\R\to\reals^\Y$.)

@@

For finite properties (those with $|\R|<\infty$) and discrete losses, we will use lowercase notation $\gamma$ and $\ell$, respectively, with reports $r\in\R$; for surrogate properties and losses we use $\Gamma$ and $L$, with reports $u\in\reals^d$.
% For finite properties (those with $|\R|<\infty$) and discrete losses, we will use lowercase notation $\gamma:\simplex\toto\R$ and $\ell:\R\toto\reals^\Y_+$, respectively, with reports $r\in\R$; for surrogate properties and losses we use $\Gamma:\simplex\toto\reals^d$ and $L:\reals^d\toto\reals^\Y_+$, with reports $u\in\reals^d$.
For general properties and losses, we will also use $\Gamma$ and $L$, as above.

@@

In fact, it is not always necessary to map \emph{all} reports to such points.
It turns out to often be convenient to allow $\ell$ to have reports that are ``redundant'' in some sense (explored in \S~\ref{sec:min-rep-sets}), initially explored by~\citet{wang2020weston} under the name \emph{embedding cardinality}.
%\bo{perhaps mention and cite \citet{wang2020weston}}
Because of this redundancy, we will only require an embedding map to be defined on a \emph{representative set}: a set of reports $\Sc$ such that, for all label distributions, at least one report $r\in\Sc$ minimizes expected loss.
\begin{definition}[Representative set]
  Let $\Gamma:\simplex\toto\R$.
  We say $\Sc \subseteq \R$ is \emph{representative for $\Gamma$} if we have $\Gamma(p) \cap \Sc \neq \emptyset$ for all $p\in \simplex$.
  We further say $\Sc$ is a \emph{minimum representative set} if it has the smallest cardinality among all representative sets.
  Given a loss $L:\R\to\reals^\Y_+$, we say $\Sc$ is a (minimum) representative set for $L$ if it is a (minimum) representative set for $\prop L$.
  % We say $\Sc \subseteq \R$ is \emph{representative for $L$} if we have $\prop{L}(p) \cap \Sc \neq \emptyset$ for all $p\in \simplex$.
  % We further say $\Sc$ is a \emph{minimum representative set} if it has the smallest cardinality among all representative sets.
\end{definition}

@@

To illustrate the idea of embedding, let us examine hinge loss in detail as a surrogate for 0-1 loss for binary classification.
Recall that we have $\R = \Y = \{-1, +1\}$, with $\hinge(u)_y = (1 - uy)_+$ and $\ellzo(r)_y := \Ind{r\neq y}$, typically with link function $\psi(u) = \sgn(u)$.
We will see that hinge loss embeds (2 times) 0-1 loss, via the embedding $\varphi(r) = r$.
For condition (i), it is straightforward to check that $\hinge(r)_y = 2\ellzo(r)_y$ for all $r,y\in\{-1,1\}$.
% First, consider that
% \begin{align*}
% 	\ellzo(-1) = (0,1) && \ellzo(1) = (1,0)\\
% 	L_{hinge}(-1) = (0,2) && L_{hinge}(1) = (2,0)\\
% \end{align*}
% Therefore, $L_{hinge}$ is twice $\ellzo$ for each $r \in \R$.
For condition (ii), let us compute the property each loss elicits, i.e., the set of optimal reports for each $p$:
\[
\prop{\ellzo}(p) = \begin{cases}
1 & p_1 > 1/2 \\
\{-1,1\} & p_1 = 1/2\\
-1 & p_1 < 1/2
\end{cases}
\qquad
\prop{L_{hinge}}(p) = \begin{cases}
[1,\infty) & p_1 = 1\\
1 & p_1 \in (1/2,1) \\
[-1,1] & p_1 = 1/2\\
-1& p_1 \in (0, 1/2)\\
(-\infty, -1]& p_1 = 0
\end{cases}~.
\]
% Now, take the embedding $\varphi$ to be the identity.  % := r$ if $r \in [-1,1]$, and $\sgn(r)$ otherwise.
In particular, we see that $-1 \in \prop{\ellzo}(p) \iff p_1 \in [0, 1/2] \iff -1 \in \prop{\hinge}(p)$, and $1 \in \prop{\ellzo}(p) \iff p_1 \in [1/2,1] \iff 1 \in \prop{\hinge}(p)$.
With both conditions of Definition~\ref{def:loss-embed} satisfied, we conclude that $\hinge$ embeds $2\ellzo$.
\raf{Forward ref \S~\ref{sec:min-rep-sets} <-- add ref to 6.2}

\begin{figure}
	\begin{minipage}{0.3\linewidth}
	\centering
	\includegraphics[width=0.95\linewidth]{figs/0-1-br.pdf}
%		\caption{Bayes risk of the 0-1 loss.}
%		\label{fig:0-1-br}
	\end{minipage}
	\hfill
	\begin{minipage}{0.3\linewidth}
	\centering		\includegraphics[width=0.95\linewidth]{figs/hinge-br.pdf}
%		\caption{Bayes risk of hinge loss.}
%		\label{fig:hinge-br}
	\end{minipage}
	\hfill
	\begin{minipage}{0.3\linewidth}
	\centering
	\includegraphics[width=0.95\linewidth]{figs/logistic-br.pdf}
%	\caption{Bayes risk of the logistic loss.}
%	\label{fig:logistic-br}
\end{minipage}
\caption{Bayes risks of 0-1, hinge, and logistic losses, respectively, in binary classification setting. \jessie{Move this down to when we actually discuss risks matching?}\raf{Let's keep but add detail / increase font size}}
\label{fig:bayes-risks-01}
\end{figure}

@@

In this section, we establish a tight relationship between the technique of embedding and the use of polyhedral (piecewise-linear convex) surrogate losses.
We defer the question of when such surrogates are consistent to \S~\ref{sec:calibration}. 
\raf{Reference Theorem~\ref{thm:embed-poly-informal}}

@@

A first observation is that a loss $L$ restricted to some representative set $\Sc$, denoted $L|_\Sc$, elicits the same property restricted to $\Sc$.
As a consequence, restricting to representative sets preserves the Bayes risk.
We will use these observations throughout.
\begin{lemma}\label{lem:loss-restrict}
  Let $L:\R\to\reals^\Y_+$ elicit $\Gamma$, and let $\Sc\subseteq\R$ be representative for $L$.
  % such that $\Gamma(p) \cap \Sc \neq \emptyset$ for all $p\in\simplex$.
  Then $L|_\Sc$ elicits $\gamma:\simplex\toto\Sc$ defined by $\gamma(p) = \Gamma(p)\cap \Sc$.
  Moreover, $\risk{L}=\risk{L|_\Sc}$.
\end{lemma}
\begin{proof}
  Let $p\in\simplex$ be fixed throughout.
  First let $r \in \gamma(p) = \Gamma(p) \cap \Sc$.
  Then $r \in \Gamma(p) = \argmin_{u\in\R} \inprod{p}{L(u)}$, so as $r\in\Sc$ we have in particular $r \in \argmin_{u\in\Sc} \inprod{p}{L(u)}$.
  For the other direction, suppose $r \in \argmin_{u\in\Sc} \inprod{p}{L(u)}$.
  As $\Sc$ is representative for $L$, we must have some $s \in \Gamma(p) \cap \Sc$.
  On the one hand, $s\in\Gamma(p) = \argmin_{u\in\R} \inprod{p}{L(u)}$.
  On the other, as $s \in \Sc$, we certainly have $s \in \argmin_{u\in\Sc} \inprod{p}{L(u)}$.
  But now we must have $\inprod{p}{L(r)} = \inprod{p}{L(s)}$, and thus $r \in \argmin_{u\in\R} \inprod{p}{L(u)} = \Gamma(p)$ as well.
  We now see $r \in \Gamma(p) \cap \Sc$.
  Finally, the equality of the Bayes risks $\min_{u\in\R} \inprod{p}{L(u)} = \min_{u\in\Sc} \inprod{p}{L(u)}$ follows immediately by the above, as $\emptyset \neq \Gamma(p)\cap\Sc \subseteq \Gamma(p)$ for all $p\in\simplex$.
\end{proof}

@@

\raf{Reference Theorem~\ref{thm:link-informal}}

@@

\raf{TODO: talk about generic approach: restrict to affine regions, take vertices}

@@

In light of this negative result, together with the fact that every polyhedral surrogate embeds some target loss (Theorem~\ref{thm:poly-embeds-discrete}), we now turn to the question: what target problem, then, \emph{is} the Lov\'asz hinge consistent for?
We show in \S~\ref{app:lovasz} that the answer is a variant of abstain loss where one can abstain on a set of labels $B\subseteq[k]$.
Recall that the symmetric difference is defined by $S\triangle T = (S\setminus T) \cup (T\setminus S)$.
\begin{theorem}\label{thm:lovasz-embed-main}
	Let $\hat\ell^f:\hat\R\times\Y\to\reals_+$, where $\hat\R = \{(A,B) \in 2^{[k]} \times 2^{[k]}: A\cap B=\emptyset\}$, be given by
	\begin{equation}
	\hat\ell^f((A,B))_S = f(A\triangle S\setminus B) + f(A\triangle S\cup B)~.
	\end{equation}
	Then the Lov\'asz hinge $L^f$ embeds $\hat\ell^f$.
  Moreover, the embedding is tight whenever $f$ is strictly submodular.
\end{theorem}
\raf{Something about how this could actually be a useful target and surrogate, provided you actually link from the surrogate to the target and not ignore $B$.}
